---
title: "Bivariate Analysis"
author: "Ian Bachli"
date: "11/7/2019"
output: html_document
---

## Setup

Here is a script that will run a bivariate analaysis of the Dengue dataset to give us a deeper look and produce additional and more complex data visualizations. This script operates using the _colmozzie_ file produced via running the _DengueProcessing_cleaning_ script located inside the _code_ folder and _processing_code_ subfolder loacted within this project.

To begin, load the required Libraries:
```{r load libaries}
library(readxl)
library(dplyr)
library(tidyverse)
library(forcats)
library(ggthemes)
library(plotly)
library(knitr)
library(naniar)
library(broom)
library(gridExtra)
library(zoo)
library(ggpubr)
```

Load the data and take a look at the variables.
```{r load data}
colmozzie <- readRDS("../../data/processed_data/processeddata.rds")

glimpse(colmozzie)
```

## Bivariate Visualization 

Bivariate Analysis is the simultaneous analysis of two variables (attributes). It explores the concept of relationship between two variables, whether there exists an association and the strength of this association, or whether there are differences between two variables and the significance of these differences. For this analysis, we will examine a few variables against each other: Average temperature of the week (TEM), Relative humidity (H), Precipitation Level (PP), and Mean wind speed (V).

```{r Comparisons}
Temp_vs_Humid <- ggplot(colmozzie, aes(x = TEM, y = H, color = TEM)) + geom_jitter()
Temp_vs_Humid

Temp_vs_Precip <- ggplot(colmozzie, aes(x = TEM, y = PP, color = PP)) + geom_jitter()
Temp_vs_Precip

Temp_vs_MeanWS <- ggplot(colmozzie, aes(x = TEM, y = V, color = V)) + geom_jitter()
Temp_vs_MeanWS

Humid_vs_Precip <- ggplot(colmozzie, aes(x = H, y = PP, color = PP)) + geom_jitter()
Humid_vs_Precip

Humid_vs_MeanWS <- ggplot(colmozzie, aes(x = H, y = V, color = V)) + geom_jitter()
Humid_vs_MeanWS

Precip_vs_MeanWS <- ggplot(colmozzie, aes(x = PP, y = V, color = V)) + geom_jitter()
Precip_vs_MeanWS
```

From what we can see, we can make some general assumptions based on the graphs as well as what we know to be common knowledge about weather effects. Looking at the graphs, Lower Temperatures are usually correlated with more more Percipitation, and Higher Humidity is correlated with higher Percipitation. While the other graphs do not seem to have much correlation.

## Seasonal Changes

Another aspect I want to look at would be examining seasonal changes and seeing as how they compare to the number of Cases of Dengue. Using this we can hopefully find some additional correlations and break down the information to get a more in depth visualization of how the number of cases of Dengue changes seasonally.

We will start with monthly data to view seasons.

```{r Seasonal Temperature Changes}
Year_vs_Temp <- ggplot(colmozzie, aes(x = Year, y = TEM, color = TEM)) + geom_jitter(alpha = 0.25) + scale_y_continuous(limits=c(20, 40)) + geom_smooth(se = FALSE) + xlab("Year") + ylab("Average Weekly Temperature (C)") + labs(color = "TEM")
Year_vs_Temp

WeekCont_vs_Temp <- ggplot(colmozzie, aes(x = Week_Continuous, y = TEM, color = TEM)) + geom_jitter(alpha = 0.25) + scale_y_continuous(limits=c(20, 40)) + geom_smooth(se = FALSE) + xlab("Weeks From 2009 to 2014") + ylab("Average Weekly Temperature (C)") + labs(color = "TEM")
WeekCont_vs_Temp

Week_vs_Temp <- ggplot(colmozzie, aes(x = Week, y = TEM, color = TEM)) + geom_jitter(alpha = 0.25) + scale_y_continuous(limits=c(20, 40)) + geom_smooth(se = FALSE) + xlab("Weeks of the Year") + ylab("Average Weekly Temperature (C)") + labs(color = "TEM")
Week_vs_Temp
```

Our temperatures seem to fluctuate normally throughout the months and years present. Now let's check out some other variables in a comparison before finally comparing it to the Cases of Dengue. What we are looking for is something interesting to account for months with higher cases of Dengue. Here we will check out how the Precipitation changes over time. 

```{r Seasonal Precipitation Changes}
Year_vs_Precip <- ggplot(colmozzie, aes(x = Year, y = PP, color = PP)) + geom_jitter(alpha = 0.25) + scale_y_continuous(limits=c(0, 50)) + geom_smooth(se = FALSE) + xlab("Year") + ylab("Average Weekly Precipitation") + labs(color = "PP")
Year_vs_Precip

WeekCont_vs_Precip <- ggplot(colmozzie, aes(x = Week_Continuous, y = PP, color = PP)) + geom_jitter(alpha = 0.25) + scale_y_continuous(limits=c(0, 50)) + geom_smooth(se = FALSE) + xlab("Weeks From 2009 to 2014") + ylab("Average Weekly Precipitation") + labs(color = "PP")
WeekCont_vs_Prercip

Week_vs_Precip <- ggplot(colmozzie, aes(x = Week, y = PP, color = PP)) + geom_jitter(alpha = 0.25) + scale_y_continuous(limits=c(0, 50)) + geom_smooth(se = FALSE) + xlab("Weeks of the Year") + ylab("Average Weekly Precipitation") + labs(color = "PP")
Week_vs_Precip
```

Now let's check out Humidity.

```{r Seasonal Humidity Changes}
Year_vs_Humid <- ggplot(colmozzie, aes(x = Year, y = H, color = H)) + geom_jitter(alpha = 0.25) + scale_y_continuous(limits=c(50, 100)) + geom_smooth(se = FALSE) + xlab("Year") + ylab("Average Weekly Humidity") + labs(color = "PP")
Year_vs_Humid

WeekCont_vs_Humid <- ggplot(colmozzie, aes(x = Week_Continuous, y = H, color = H)) + geom_jitter(alpha = 0.25) + scale_y_continuous(limits=c(50, 100)) + geom_smooth(se = FALSE) + xlab("Weeks From 2009 to 2014") + ylab("Average Weekly Humidity") + labs(color = "PP")
WeekCont_vs_Humid

Week_vs_Humid <- ggplot(colmozzie, aes(x = Week, y = H, color = H)) + geom_jitter(alpha = 0.25) + scale_y_continuous(limits=c(50, 100)) + geom_smooth(se = FALSE) + xlab("Weeks of the Year") + ylab("Average Weekly Humidity") + labs(color = "PP")
Week_vs_Humid
```

Next, Windspeed. Windspeed might be an odd one to examine, but I am interested in looking at how Windspeed might play a role as a deterant for vectors.

```{r Seasonal Windspeed Changes}
Year_vs_Windspeed <- ggplot(colmozzie, aes(x = Year, y = V, color = V)) + geom_jitter(alpha = 0.25) + scale_y_continuous(limits=c(0, 30)) + geom_smooth(se = FALSE) + xlab("Year") + ylab("Average Weekly Windspeed") + labs(color = "PP")
Year_vs_Windspeed

WeekCont_vs_Windspeed <- ggplot(colmozzie, aes(x = Week_Continuous, y = V, color = V)) + geom_jitter(alpha = 0.25) + scale_y_continuous(limits=c(0, 30)) + geom_smooth(se = FALSE) + xlab("Weeks From 2009 to 2014") + ylab("Average Weekly Windspeed") + labs(color = "PP")
WeekCont_vs_Windspeed

Week_vs_Windspeed <- ggplot(colmozzie, aes(x = Week, y = V, color = V)) + geom_jitter(alpha = 0.25) + scale_y_continuous(limits=c(0, 30)) + geom_smooth(se = FALSE) + xlab("Weeks of the Year") + ylab("Average Weekly HWindspeed") + labs(color = "PP")
Week_vs_Windspeed
```

So far, everything seems to line up rather well with Seasonal expectancies. 


